---
title: "Example with random forest model"
author: "Anna Kozak"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Example with random forest regression model

In this vignette we present measure for random forest regression model.
```{r, warning = FALSE, echo = FALSE, message = FALSE, include = TRUE}
library(LocalVariableImportanceViaOscillations)
library(DALEX)
library(randomForest)
library(ingredients)
```

```{r, include=FALSE, warning=FALSE, error=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
```
### Dataset
We work on Apartments dataset.

```{r, warning = FALSE, echo = FALSE, message = FALSE, include = TRUE}
data(apartments)
head(apartments)
```

### Ranfom forest regression model

Now, we define a model and use explain from `DALEX`.

```{r, warning = FALSE, error = FALSE, message = FALSE, include = TRUE}
apartments_rf_model <- randomForest(m2.price ~ construction.year + surface + floor +
                                      no.rooms, data = apartments)
explainer_rf <- explain(apartments_rf_model,
                        data = apartmentsTest[,2:5], y = apartmentsTest$m2.price)
```

We need to specify an observation. Let consider a new apartment with following attributes. 

```{r, warning = FALSE, error = FALSE, message = FALSE, include = TRUE}
new_apartment <- data.frame(construction.year = 1998, surface = 88, floor = 2L, no.rooms = 3)
cp <- ingredients::ceteris_paribus(explainer_rf, new_apartment)
```


Let see the Ceteris Paribus Plots calculted with `ceteris_paribus()` function.

```{r, warning = FALSE, error = FALSE, message = FALSE, include = TRUE}
plot(cp) + show_observations(cp)
```

Now, we calculated a measure of local variable importance via oscillation based on Ceteris Paribus plot.

```{r, warning = FALSE, error = FALSE, message = FALSE, include = TRUE}
measure <- LocalVariableImportance(cp, apartments[,2:5], absolute_deviation = TRUE, point = TRUE, density = TRUE)
```

```{r, warning = FALSE, error = FALSE, message = FALSE, include = TRUE}
plot(measure)
```
